---
title: "Seminario de Investigación"
format:
  revealjs:
    theme: custom.scss
    width: 1600
    height: 900
    navigation-mode: vertical
    controls-layout: edges
    controls-tutorial: true
    callout-appearance: default
    slide-number: c/t
    footer: "Seminario de Investigación — Seminar (RED)"
    logo: "logo.png"
    self-contained: true
    chalkboard: false
project:
  type: website
  output-dir: public
bibliography: bibliord.bib
---

#  {text-align="center" data-menu-title="Introduction"}

```{r}
#| echo: false
library(tidyverse)
library(fontawesome)
library(crayon)
library(knitr)
library(ggthemes)
library(kableExtra)
library(directlabels)
library(ggtext)
library(plotly)
library(showtext)
library(countrycode)
library(readxl)
library(gt)
library(gtExtras)
library(countdown)
library(here)

knitr::knit_hooks$set(crop = knitr::hook_pdfcrop)
```

<br><br><br><br>

::: title
Seminario de Investigación
:::

::: subtitle
Seminar 1: Plagiarism, fishing, and the limits of the scientific method
:::

::: author
Sergi Martínez
:::

<br><br><br><br>

::: media
`r fontawesome::fa(name = "envelope", fill = "#2A363B")` [smsoler\@clio.uc3m.es]{style="color:#1DA1F2"} \| `r fontawesome::fa(name = "map-pin", fill = "#2A363B")` [Room 6.01.02]{style="color:#2A363B"}
:::

## How seminars work {.scrollable}

::: small

- 11 sessions across the semester, combining:

  - Critical discussion of the assigned readings

  - Presentation of your group project

  - Applied practice in R

:::

. . .

Two components of your grade come directly from this room:

::: small

- **Group critical presentation** (10%): each group presents and critically evaluates one assigned article, in the week it's assigned — 15'

- **Three critical questions** (10%): across the semester, each of you must raise at least three critically informed questions about the papers or the seminar content

:::

## Today's readings {.scrollable}

::: small

- Don Green / Michael LaCour episode — two press pieces:

  1. PBS NewsHour, *"An Author Of Paper About Changing Minds On Gay Marriage Seeks Retraction"* (2015)

  2. Retraction Watch, *"10 years after the downfall of a same-sex marriage canvassing study"* (2025)

- Broockman, D., & Kalla, J. (2016). Durably reducing transphobia: A field experiment on door-to-door canvassing. *Science*, 352(6282), 220–224.

:::

## The finding that made headlines {.scrollable}

LaCour & Green (2014), *"When Contact Changes Minds"*, published in **Science**:

- A single ~20-minute doorstep conversation with a gay canvasser durably shifted people's support for same-sex marriage

. . .

- Covered by *This American Life*, the *NYT*, the *WSJ*, the *Washington Post*, *Vox*, and more — a dream result, and a dream case for teaching field experiments

. . .

- Two Berkeley graduate students, David Broockman and Joshua Kalla, admired the design enough to want to extend it themselves

## The unraveling {.scrollable}

::: small

While preparing their own extension, Broockman, Kalla, and Peter Aronow (Yale) tried to reconstruct LaCour's dataset — and kept hitting inconsistencies

:::

. . .

- They documented **eight statistical irregularities**, none conclusive alone, but jointly pointing the same way

- Their report's conclusion, in careful academic language: the data *"were not collected as described"*

. . .

<br>

- They brought it to Donald Green, the senior co-author, who confronted LaCour — who confessed to fabricating the data

- Green requested the retraction himself; *Science* retracted the paper in May 2015

## Discuss {.discussion .center}
```{css}
.center h2 {
  text-align: center;
}
```

Why didn't peer review catch this? What made the finding so attractive to fabricate in the first place? What did *replication* — not review — end up catching?

```{r}
#| echo: false
countdown::countdown(minutes = 8, warn_when = 60)
```

## From scandal to science {.scrollable}

Retracting a paper doesn't answer the underlying question — it just leaves it open again.

. . .

- Broockman & Kalla went on to run the study properly themselves: real door-to-door canvassing in Miami, this time on **transphobia**, with a placebo topic (recycling) as the control conversation

- Result: genuine, durable reductions in prejudice — replicated later in other contexts

. . .

<br>

*"Science advances incrementally"* — a fabricated shortcut doesn't survive; a slow, careful, replicable design does.

## Fishing expeditions {.scrollable}

Fabrication is the extreme case. Far more common, and much harder to spot, is **fishing**:

- Multiple comparisons — testing many outcomes, subgroups, or specifications until one is significant

- The garden of forking paths — the same dataset can support many "reasonable" analysis choices; which one gets published?

- HARKing — Hypothesizing After the Results are Known, then writing the paper as if the hypothesis came first

. . .

Each of these inflates the rate of false positives, without anyone needing to fabricate a single number.

## The limits of "the" scientific method, applied to social reality {.scrollable}

::: small

Political and social phenomena resist the clean version of the scientific method you may associate with a lab:

:::

- Most treatments of interest can't be randomly assigned (a war, an institution, an identity)

- Confounding is the default, not the exception, in observational data

- Context matters: an effect found in one setting may not travel to another (external validity)

- The subjects are reflexive — people react to being studied, to incentives, to the researcher

. . .

This is part of why the 2010s brought a **replication crisis** across psychology, economics, and political science — and why our field now leans harder on preregistration, transparency, and replication.

## Safeguards for this semester {.scrollable}

::: small

- Cite properly — every idea or text you use must be attributed (APA, for instance)

- Share your data and code — reproducibility is the point of the R workshops later this semester

- AI tools are a powerful aid for editing and clarity — **not** for writing from scratch or inventing references

:::

. . .

<br>

**Any hallucinated citations or evidence, or any form of plagiarism — AI-generated or otherwise — is evaluated with a grade of zero.**

## Group formation

Since our very first scheduled seminar (Sept 7) was cancelled, we'll use today's session to:

1. Form your project groups

2. Set the rota for the group critical presentations across the semester

## Looking ahead

**Friday, Sept 18 — Lecture 2**: The research process I: from the question to hypotheses.

<br>

**Monday, Sept 21 — Seminar**: Motivation and research questions. Read Cyrus Samii's blog and identify three societal problems — we'll discuss how to turn them into research questions.

## See you Friday! smsoler\@clio.uc3m.es {.center}
```{css}
.center h2 {
  text-align: center;
}
```
